---
title: "Statistical Inference Course Project Part 2"
author: "Yan"
date: "5/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 2: Basic Inferential Data Analysis Instructionsless 

Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

#### Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
library(datasets)
library(ggplot2)
data(ToothGrowth)
head(ToothGrowth)
```

```{r}
tg <- ToothGrowth
tg$dose <- as.factor(tg$dose)
qplot(x=supp,y=len,data = tg,facets = .~dose) + geom_boxplot(aes(fill = supp))
```

#### Provide a basic summary of the data.

```{r}
dim(tg)
summary(tg)
```

#### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 

Note: (Only use the techniques from class, even if there's other approaches worth considering)

split the dataset according to dose and supp
```{r}
OJ <- tg[tg$supp == "OJ",]
VC <- tg[tg$supp == "VC",]
OJdose1 <- tg[tg$supp == "OJ" & tg$dose == 0.5,]
OJdose2 <- tg[tg$supp == "OJ" & tg$dose == 1,]
OJdose3 <- tg[tg$supp == "OJ" & tg$dose == 2,]
VCdose1 <- tg[tg$supp == "VC" & tg$dose == 0.5,]
VCdose2 <- tg[tg$supp == "VC" & tg$dose == 1,]
VCdose3 <- tg[tg$supp == "VC" & tg$dose == 2,]
dim(OJ)
dim(VC)
dim(OJdose1)
dim(OJdose2)
dim(OJdose3)
dim(VCdose1)
dim(VCdose2)
dim(VCdose3)
```

Check if tooth growth is affected by dose


##### Get confidence intervals for different supp regardless different dose

```{r}
t.test(len ~ supp, data = tg,paired=FALSE, var.equal = FALSE )$conf
```

So when doesn't consider the dosage, doesn't show significant difference between two supplements.

##### Get confidence intervals for low/median/large dose with different supp

```{r}
t.test(OJdose1$len, VCdose1$len, paired = FALSE, var.equal = FALSE)$conf
t.test(OJdose2$len, VCdose2$len, paired = FALSE, var.equal = FALSE)$conf
t.test(OJdose3$len, VCdose3$len, paired = FALSE, var.equal = FALSE)$conf
```

From the 95% confidence intervals, 0 is not included when the dosage is small or median. So there's significant difference between two supplement with OJ leads to faster tooth growth than VC.

##### Get confidence intervals for same supp with different dosage

```{r}
t.test(OJdose1$len, OJdose2$len, paired = FALSE, var.equal = FALSE)$conf
t.test(OJdose1$len, OJdose3$len, paired = FALSE, var.equal = FALSE)$conf
t.test(OJdose2$len, OJdose3$len, paired = FALSE, var.equal = FALSE)$conf
t.test(VCdose1$len, VCdose2$len, paired = FALSE, var.equal = FALSE)$conf
t.test(VCdose1$len, VCdose3$len, paired = FALSE, var.equal = FALSE)$conf
t.test(VCdose2$len, VCdose3$len, paired = FALSE, var.equal = FALSE)$conf
```

From the 95% confidence intervals, 0 is not included in all cases. So there's significant difference between different dosage of same supplement, the tooth growth becomes faster with increasing dosage.

##### State your conclusions and the assumptions needed for your conclusions.

Assumptions:  

* The varience with different supplement and different dosage is not the same.

* The data between different groups are not paired

* The sample population is not skewed

Conclusions:  

* With same supplement, OJ or VC, increasing the dosage leads to faster tooth growth

* With small or median dosage, OJ leads to faster tooth growth than VC

* With large dosage, no significant difference observed using OJ or VC



## Review criteria (self check)  

* [x] Did you show where the distribution is centered at and compare it to the theoretical center of the distribution?

* [x] Did you show how variable it is and compare it to the theoretical variance of the distribution?

* [x] Did you perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?

* [x] Did the student perform some relevant confidence intervals and/or tests?

* [x] Were the results of the tests and/or intervals interpreted in the context of the problem correctly?

* [x] Did the student describe the assumptions needed for their conclusions?


##### Links used:  

[link 1](https://rpubs.com/gviglioni/ExponentialDistribution_CentralLimitTheorem)  
[link 2](https://rpubs.com/a_inparadise/86939)  
[link 3](https://rpubs.com/abhiuee/53828)  
[link 4](https://rstudio-pubs-static.s3.amazonaws.com/82410_9df3ec9e695c45eda138f47a2fae9187.html)  
[link 5](http://rstudio-pubs-static.s3.amazonaws.com/156585_ea5571a742cc4f1ea94e401a7320c4a4.html)  
[link 6](https://rstudio-pubs-static.s3.amazonaws.com/88003_457d504772954be399c4070d9b404a0f.html)  
[link 7](https://rstudio-pubs-static.s3.amazonaws.com/30490_f5f75fae348843819993bf1e0957f4f1.html)  
[link 8](https://rstudio-pubs-static.s3.amazonaws.com/67663_669e39c2153e4f8295f9d130abf07b7e.html)  

